---
title: "regression_code"
author: "Elle Strauss, Ananya Sista"
date: "`r Sys.Date()`"
output: pdf_document
---
# Package Installations
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(MASS)      # LDA + stepAIC
library(rpart)     # decision trees
library(caret)     # evaluation metrics
library(pROC)      # ROC / AUC
library(ggplot2)
library(rpart.plot)
```

# Cleaning Data
```{r}
# read in csv (make sure working directory is the repo)
oscar_noms <- read_csv('all_tmbd_movie_info.csv')
colnames(oscar_noms) <- c('year', 'ceremony', 'award', 'winner', 'film_name', 'film_producers', 'revenue', 'budget', 'runtime', 'vote_average_list', 'vote_count', 'popularity', 'genre', 'imdb_rating', 'rotten_tomato_score', 'mppa_rating', 'directors')

# cleaning data (removing any movies that didn't have a budget or revenue)
# ended up removing 20 observations
oscars_bp <- oscar_noms %>% filter(revenue > 0 & budget > 0)
# do log of revenue and budget (highly skewed to the right), not linear
oscars_bp$log_revenue <- log(oscars_bp$revenue)
oscars_bp$log_budget <- log(oscars_bp$budget)
oscars_bp$ROI <- oscars_bp$revenue / oscars_bp$budget

summary(oscars_bp)

# confirm winner is a numeric
oscars_bp$winner <- factor(oscars_bp$winner)
```

## Separating Out Critic, Audience, and Market Variables
```{r}
critics_vars <- c("rotten_tomato_score")

audience_vars <- c("imdb_rating", "vote_average_list", "vote_count")

market_vars <- c("log_revenue", "log_budget", "popularity", "ROI")
```

# Logistic Regression
## Individual Predictors
```{r}
AIC1.1 <- matrix(rep(0,9), ncol=1)
rownames(AIC1.1) <- c("log_revenue", "log_budget", "runtime", "vote_average_list", "vote_count", "popularity", "imdb_rating", "rotten_tomato_score", "ROI")

mod1.1a <- glm(winner ~ log_revenue, family=binomial, data=oscars_bp)
summary(mod1.1a)
AIC1.1[1] <- AIC(mod1.1a)
predict1.1a <- predict(mod1.1a, type="response")
oscars1.1a <- ifelse(predict1.1a >= 0.5,1,0)
table(oscars_bp$winner, oscars1.1a)

mod1.1b <- glm(winner ~ log_budget, family=binomial, data=oscars_bp)
summary(mod1.1b)
AIC1.1[2] <- AIC(mod1.1b)
predict1.1b <- predict(mod1.1b, type="response")
oscars1.1b <- ifelse(predict1.1b >= 0.5,1,0)
table(oscars_bp$winner, oscars1.1b)

mod1.1c <- glm(winner ~ runtime, family=binomial, data=oscars_bp)
summary(mod1.1c)
AIC1.1[3] <- AIC(mod1.1c)
predict1.1c <- predict(mod1.1c, type="response")
oscars1.1c <- ifelse(predict1.1c >= 0.5,1,0)
table(oscars_bp$winner, oscars1.1c)

mod1.1d <- glm(winner ~ vote_average_list, family=binomial, data=oscars_bp)
summary(mod1.1d)
AIC1.1[4] <- AIC(mod1.1d)
predict1.1d <- predict(mod1.1d, type="response")
oscars1.1d <- ifelse(predict1.1d >= 0.5,1,0)
table(oscars_bp$winner, oscars1.1d)

mod1.1e <- glm(winner ~ vote_count, family=binomial, data=oscars_bp)
summary(mod1.1e)
AIC1.1[5] <- AIC(mod1.1e)
predict1.1e <- predict(mod1.1e, type="response")
oscars1.1e <- ifelse(predict1.1e >= 0.5,1,0)
table(oscars_bp$winner, oscars1.1e)

mod1.1f <- glm(winner ~ popularity, family=binomial, data=oscars_bp)
summary(mod1.1f)
AIC1.1[6] <- AIC(mod1.1f)
predict1.1f <- predict(mod1.1f, type="response")
oscars1.1f <- ifelse(predict1.1f >= 0.5,1,0)
table(oscars_bp$winner, oscars1.1f)

mod1.1g <- glm(winner ~ imdb_rating, family=binomial, data=oscars_bp)
summary(mod1.1g)
AIC1.1[7] <- AIC(mod1.1g)
predict1.1g <- predict(mod1.1g, type="response")
oscars1.1g <- ifelse(predict1.1g >= 0.5,1,0)
table(oscars_bp$winner, oscars1.1g)

mod1.1h <- glm(winner ~ rotten_tomato_score, family=binomial, data=oscars_bp)
summary(mod1.1h)
AIC1.1[8] <- AIC(mod1.1h)
predict1.1h <- predict(mod1.1h, type="response")
oscars1.1h <- ifelse(predict1.1h >= 0.5,1,0)
table(oscars_bp$winner, oscars1.1h)


mod1.1i <- glm(winner ~ ROI, family=binomial, data=oscars_bp)
summary(mod1.1i)
AIC1.1[9] <- AIC(mod1.1i)
predict1.1i <- predict(mod1.1i, type="response")
oscars1.1i <- ifelse(predict1.1i >= 0.5,1,0)
table(oscars_bp$winner, oscars1.1i)

AIC1.1
```

### Plots
```{r}
aic_df <- data.frame(
  predictor = rownames(AIC1.1),
  AIC = AIC1.1[,1]
)

plot(
  aic_df$AIC,
  type = "b",
  xaxt = "n",
  xlab = "Predictor",
  ylab = "AIC",
  main = "AIC by Single-Predictor Logistic Models"
)

axis(1, at = 1:nrow(aic_df), labels = aic_df$predictor, las = 2)

ggplot(aic_df, aes(x = reorder(predictor, AIC), y = AIC)) +
  geom_point(size = 3) +
  geom_line(group = 1) +
  labs(
    title = "AIC by Single-Predictor Logistic Models",
    x = "Predictor",
    y = "AIC"
  )
```

## Comparing Models (All Predictors)
```{r}
mod.full <- glm(winner ~ log_revenue + log_budget + runtime + vote_average_list + vote_count + popularity + imdb_rating + rotten_tomato_score + ROI, family=binomial, data=oscars_bp)
mod.null <- glm(winner ~ 1, family=binomial, data=oscars_bp) # intercept only model

stepAIC(mod.full, direction="backward")
# best backward model: log_revenue + log_budget + runtime + vote_average_list + popularity
stepAIC(mod.null, direction="forward", scope=list(upper=mod.full, lower=mod.null))
# best forward model: imdb_rating + ROI + runtime + log_budget + log_revenue
```

## Predicting Based on Best Fit
### Adding Test and Train Data
```{r}
set.seed(123)

train_idx <- sample(1:nrow(oscars_bp), 0.7 * nrow(oscars_bp))

train <- oscars_bp[train_idx, ]
test  <- oscars_bp[-train_idx, ]
```


```{r}
# Try out the best fit on predicting (use backwards)
bestFitBack <- glm(winner ~ log_revenue + log_budget + runtime + vote_average_list + popularity, family=binomial, data=train)
pred_prob_back <- predict(bestFitBack, newdata = test, type = "response")
pred_back <- ifelse(pred_prob_back >= 0.5, 1, 0)

table(Actual = test$winner, Predicted = pred_back)

# oscars_bp$predict_BestBack <- predict(bestFitBack, type="response")
# oscars.1BestBack <- oscars_bp %>% group_by(year) %>% mutate(predicted = 0, predicted = replace(predicted, which.max(predict_BestBack), 1)) %>% ungroup()
# table(oscars.1BestBack$winner, oscars.1BestBack$predicted)

bestFitFor <- glm(winner ~ imdb_rating + ROI + runtime + log_budget + log_revenue, family=binomial, data=train)
pred_prob_for <- predict(bestFitFor, newdata = test, type = "response")
pred_for <- ifelse(pred_prob_for >= 0.5, 1, 0)

table(Actual = test$winner, Predicted = pred_for)
# oscars_bp$predict_BestFor <- predict(bestFitFor, type="response")
# oscars.1BestFor <- oscars_bp %>% group_by(year) %>% mutate(predicted = 0, predicted = replace(predicted, which.max(predict_BestFor), 1)) %>% ungroup()
# table(oscars.1BestBack$winner, oscars.1BestFor$predicted)
```
## Model Comparisons
```{r}
# Critics only model
model_critics <- glm(
  winner ~ rotten_tomato_score,
  family = binomial,
  data = train
)

# Audience only model
model_audience <- glm(
  winner ~ imdb_rating + vote_average_list + vote_count,
  family = binomial,
  data = train
)

# Combined model (critics + audience)
model_crit_aud <- glm(
  winner ~ rotten_tomato_score +
    imdb_rating + vote_average_list + vote_count,
  family = binomial,
  data = train
)

# Market model (bugdet, popularity, revenue)
model_market <- glm(winner ~ log_revenue + log_budget + popularity,
                    family = binomial, data = train)

# Full model (controls included)
model_full <- glm(
  winner ~ rotten_tomato_score +
    imdb_rating + vote_average_list + vote_count +
    log_revenue + log_budget + runtime + popularity,
  family = binomial,
  data = train
)



AIC(model_critics, model_audience, model_crit_aud, model_market, model_full)

# predicted probabilities
p_critics <- predict(model_critics, newdata = test, type = "response")
p_audience <- predict(model_audience, newdata = test, type = "response")
p_crit_aud <- predict(model_crit_aud, newdata = test, type = "response")
p_market <- predict(model_market, newdata = test, type = "response")
p_full <- predict(model_full, newdata = test, type = "response")

# ROC curves
roc_critics <- roc(test$winner, p_critics)
roc_audience <- roc(test$winner, p_audience)
roc_crit_aud <- roc(test$winner, p_crit_aud)
roc_market <- roc(test$winner, p_market)
roc_full <- roc(test$winner, p_full)

plot(roc_critics, col="red3")
lines(roc_audience, col="gold")
lines(roc_crit_aud, col="blue3")
lines(roc_market, col="lightgreen")
lines(roc_full, col="magenta")


legend("bottomright",
       legend=c("Critics", "Audience", "Combined", "Market", "Full"),
       col=c("red3","gold","blue3", "lightgreen", "magenta"),
       lwd=2)

# AUC values
auc(roc_critics)
auc(roc_audience)
auc(roc_crit_aud)
auc(roc_market)
auc(roc_full)

auc_df <- data.frame(
  model = c("Critics","Audience","Combined","Market","Full"),
  auc = c(auc(roc_critics),
          auc(roc_audience),
          auc(roc_crit_aud),
          auc(roc_market),
          auc(roc_full))
)

barplot(auc_df$auc, names.arg = auc_df$model, main="Model Comparison", ylim=c(0, 1), col="lightblue", xlab="Models", ylab="AUC")
```

### Statistical Test of Models
```{r}
anova(model_audience, model_crit_aud, test = "Chisq")
anova(model_critics, model_crit_aud, test = "Chisq")
anova(model_market, model_full, test = "Chisq")
anova(model_crit_aud, model_full, test = "Chisq")

# Predictions
pred_critics <- ifelse(p_critics >= 0.5, 1, 0)
pred_audience <- ifelse(p_audience >= 0.5, 1, 0)
pred_crit_aud <- ifelse(p_crit_aud >= 0.5, 1, 0)
pred_market <- ifelse(p_market >= 0.5, 1, 0)

# Confusion matrices
table(test$winner, pred_critics)
table(test$winner, pred_audience)
table(test$winner, pred_crit_aud)
table(test$winner, pred_market)

summary(model_crit_aud)

exp(cbind(Odds_Ratio = coef(model_crit_aud),
          confint(model_crit_aud)))
```



### Confusion Table
Left to Right  
True Negative: Movie did not win, model correctly predicted no win.  
False Positive: Movie lost, but model predicted winner.  
False Negative: Movie actually won, but model predicted loser.  
True Positive: Movie won, model correctly predicted winner.

### Comparing Step Models
```{r}
roc_back <- roc(test$winner, pred_prob_back)
roc_for  <- roc(test$winner, pred_prob_for)

plot(roc_back, col="blue", main="ROC Comparison of Logistic Models")
lines(roc_for, col="red")

legend("bottomright",
       legend=c("Backward Model", "Forward Model"),
       col=c("blue", "red"),
       lwd=2)

# Coefficients Odds Ratio
# visualizes the estimated effects of predictor variables on a categorical outcome by displaying their log-odds or odds ratios.
coef_df <- data.frame(
  variable = names(coef(bestFitBack)),
  value = coef(bestFitBack)
)

ggplot(coef_df[-1,], aes(x = reorder(variable, value), y = value)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(
    title = "Logistic Regression Coefficients",
    x = "Predictor",
    y = "Effect on Log-Odds"
  )
```

### Probability Of Separation
```{r}
ggplot(test, aes(x = pred_prob_back, fill = winner)) +
  geom_density(alpha = 0.5) +
  labs(
    title = "Separation of Predicted Probabilities",
    x = "Predicted Probability of Winning"
  )
```
Does the model actually separate winners from losers? 
If curves overlap → weak model  
If separated → strong model  

# Classification Tree
```{r}
# Classification Tree
tree_model <- rpart(winner ~ log_revenue + log_budget + imdb_rating + popularity + vote_count + runtime, data=oscars_bp, method="class")
tree_pred <- predict(tree_model, newdata = test, type="class")
table(Actual=test$winner, Predicted=tree_pred)
print(tree_model)

# Classification Tree Plot
rpart.plot(tree_model,
           type = 2,
           extra = 104,
           main = "Classification Tree for Oscar Prediction")

# Variable Importance
barplot(tree_model$variable.importance,
        main = "Variable Importance (Decision Tree)",
        las = 2, ylim=c(0, 12), col="lightblue", ylab="Variable Importance", xlab="Variable")


```


# LDA
```{r}
model <- lda(winner ~ log_revenue + log_budget + imdb_rating + popularity, data=train)

lda_pred <- predict(model, newdata=test)$class

table(Actual=test$winner, Predicted=lda_pred)
```
# General Analysis
```{r}
# Means
mean(pred_back == test$winner)
mean(lda_pred == test$winner)
mean(tree_pred == test$winner)

# Confusion Matrix
confusionMatrix(as.factor(pred_back), test$winner)
confusionMatrix(as.factor(lda_pred), test$winner)
confusionMatrix(as.factor(tree_pred), test$winner)

# ROC/AUC
# A probability curve that plots the True Positive Rate (Sensitivity)
roc_obj <- roc(test$winner, pred_prob_back)



plot(roc_obj, col="blue3")
lines(roc_lda, col="lightgreen")
lines(roc_tree, col="magenta")


legend("bottomright",
       legend=c("Backwards","LDA","Classification Tree"),
       col=c("blue3", "lightgreen", "magenta"),
       lwd=2)

# AUC values
auc_df <- data.frame(
  model = c("Backwards","LDA","Classification Tree"),
  auc = c(auc(roc_obj),
          auc(roc_lda),
          auc(roc_tree))
)

barplot(auc_df$auc, names.arg = auc_df$model, main="Model Comparison", ylim=c(0, 1), col="lightblue", xlab="Models", ylab="AUC")
```

